---
output: html_document
---
Assignment 1
==================================
##Fernando Colugnati


This is the first peer reviewed assignement for the "Reproducible Research" at Coursera, January, 2015.

##Loading and preprocessing the data
Just unziped and loaded data using simple read.cvs function. Also, packages that should be installed are listed. In some tests I could not point the mirror correctely, resulting error.

```{r load data}
#Read the file
unzip("activity.zip")
a<-read.csv("activity.csv")

#Install packages
#install.packages("plyr")
#install.packages("Hmisc")
#install.packages("doBy")

```


##What is mean total number of steps taken per day?

Fisrts table presents simple statistics for the whole dataset, and also descritptivestatistics for the total number of steps in each day.

```{r summarize data, warning=FALSE, message=FALSE}
#summarizes dataset and shows statistics for daily steps
summary(a)
desc <- summary(tapply(a$steps, a$date, sum))
ldesc <- summary(tapply(log(a$steps+1), a$date, sum))
desc
ldesc
```

The mean and median for the total daily steps are `r desc["Mean"]` and `r desc["Median"]`, respectively. Also, the log of the same variable plus one is described.

Histograms bellow presents the distribution for the total number of daily steps.

```{r ploting steps by interval}

par(mfrow = c(1, 2))
hist(tapply(a$steps, a$date, sum), main="Daily steps distribution", xlab="Total daily steps")
hist(tapply(log(a$steps+1), a$date, sum), main=" ", xlab="Log(total steps + 1)")
```


## What is the average daily activity pattern?


```{r}
c <- aggregate(steps ~ interval, data=a, FUN=mean)
attach(c)
par(mfrow = c(1, 1))
plot(c, type="l")

c <- c[order(-steps),]
```

Across all days, considering the same 5 minutes interval, it is possible to note a activity in the first intervals, maybe during the morning, reaching the maximun average number of steps at  interval `r c[1,1]`, reaching arround `r round(c[1,2])` steps.


##Imputing missing values

I am using functions from plyr and Hmisc packages, that allows to input values for NA's easily, as follows. I used two strategies, considering the interval means and medians for the number of steps.

```{r input missings, warning=FALSE, message=FALSE}
require(plyr)
require(Hmisc)

a <- ddply(a, "interval", mutate, steps.mean = impute(steps, mean))
a <- ddply(a, "interval", mutate, steps.median = impute(steps, median))


summary(a)

par(mfrow = c(1, 2))
hist(tapply(a$steps.mean, a$date, sum), main="Daily steps - mean inputed", xlab="Total daily steps")
hist(tapply(a$steps.median, a$date, sum), main="Daily steps - median inputed", xlab="Total daily steps")
```


There are `r desc("NA")` missing values in the original dataset.  Apparently, there are no big impacts regarding mean and median, but the 3rd quartile varies. As can be noted in histograms, the medians increased the number of 0 (zeroes). 

##Are there differences in activity patterns between weekdays and week- ends?

Follows how I created the factor. I have to use a setup so I can have weekdays in English format.

```{r create weekday factor}
Sys.setlocale("LC_TIME", "C") #I am in Brazil, must do this!

a$weekday <- weekdays(as.Date(a$date), abbreviate=TRUE)

a$weekday_f[a$weekday=="Sat" | a$weekday=="Sun"]<-0
a$weekday_f[a$weekday=="Mon" | a$weekday=="Tue" | a$weekday=="Wed" | 
            a$weekday=="Fri" | a$weekday=="Thu"]<-1
a$weekday_f <- factor(a$weekday_f, levels=c(0,1), labels=c("Weekends", "Weekdays"))

table(a$weekday)
table(a$weekday_f)
```

Regarding differences between weekdays and weekends, plots bellow, as well as simple statistical tests, shows that this guys walks a little bit more during weekends. In avarage or in median, he walks 7 steps more during weekends.


```{r}
a_weekend <- subset(a, weekday_f=="Weekends")
a_weekday <- subset(a, weekday_f=="Weekdays")

par(mfrow = c(2, 1))
plot(aggregate(steps.mean ~ interval, data=a_weekend, FUN=mean), type="l", sub="Weekend")
plot(aggregate(steps.mean ~ interval, data=a_weekday, FUN=mean), type="l", sub="weekday")

par(mfrow = c(1, 2))
boxplot(a$steps.mean ~ a$weekday,data=a, main="Steps Distribution" , ylab="Number of steps")
boxplot(log(steps.mean+1) ~ weekday,data=a , ylab="Log(Number of steps +1)")

par(mfrow = c(1, 2))
boxplot(a$steps~a$weekday_f,data=a, main="Steps Distribution" , ylab="Number of steps")
boxplot(log(steps+1)~weekday_f,data=a, ylab="Log(Number of steps +1)")

```

```{r}
library(doBy)
summaryBy(steps.mean + steps.median ~ weekday_f, data = a, 
          FUN = function(x) { c(m = mean(x), s = sd(x), md = median(x)) } )

wilcox.test(a_weekend$steps.mean,a_weekday$steps.mean)
wilcox.test(log(a_weekend$steps.mean+1),log(a_weekday$steps.mean+1))

t.test(a_weekend$steps.mean,a_weekday$steps.mean)
t.test(log(a_weekend$steps.mean+1),log(a_weekday$steps.mean+1))


```

Very funny and usefull assignment! Thank you, I hope I could achieve the expected performance.


